---
title: "OSHetero2021 Loading and Processing"
author: "Sanjana Rajan and Emily Franz"
date: "`r format(Sys.time(), '%m/%d/%Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 5
    number_sections: false
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  tidy = TRUE,
  echo = TRUE,
  cache = TRUE,
  collapse = TRUE,
  tidy.opts = list(width.cutoff = 95),
  message = FALSE,
  warning = FALSE,
  cache.lazy = FALSE
)
```

```{r Fig1_lib, cache=FALSE}
setwd("/gpfs0/home2/gdrobertslab/rsexf001/OSHetero2021")
source("scSeurat.R")
source("Downstream.v2.R")

# loading libraries
library(Seurat)
library(future)
library(ggplot2)
# library(fgsea)
library(msigdbr)
library(clusterProfiler)
library(SingleR)
library(dplyr) # %<%
library(pheatmap)
library(RColorBrewer)
library(viridis) # inferno color palette
library(grid)
library(ggsci)
library(data.table)
library(ggvenn)
```

# Load Seurat objects 

Load Seurat objects here so only have to complete once (commented out elsewhere). 

## Load Osteoblasts

```{r load_ob}
# Create Seurat objects and perform initial QC.  Label original source.
# OB
OB <- tenXLoadQC(path10x = "/gpfs0/home/gdrobertslab/lab/Counts/S0031/outs/filtered_feature_bc_matrix/",
                 spec = "human")

OB <- subset(OB,
             subset = nCount_RNA < 20000 &
               percent.mt < 10)

OB$src <- "OB"
OB$cond <- "OB"
```

## Load OS17

```{r load_os17}
# OS17
## Culture
os17.cx.raw <- tenXLoadQC(path10x = "/gpfs0/home/gdrobertslab/lab/Counts/S0016/outs/filtered_feature_bc_matrix/",
                          spec = "mixHuman")

os17.cx.raw <- subset(os17.cx.raw,
                      subset = nFeature_RNA > 3500 &
                        nCount_RNA < 50000 &
                        percent.mt < 15)

os17.cx.raw$src <- "Culture"
os17.cx.raw$cond <- "OS17"

## Tibia
os17.tib.raw <- tenXLoadQC(path10x = "/gpfs0/home/gdrobertslab/lab/Counts/S0018xS0028/outs/filtered_feature_bc_matrix/",
                           spec = "mixHuman")
os17.tib.raw <- subset(os17.tib.raw,
                       subset = nFeature_RNA > 3000 &
                         nCount_RNA < 60000 &
                         percent.mt < 18)

os17.tib.raw$src <- "Tibia"
os17.tib.raw$cond <- "OS17"

## Lung
## double check usage throughout rest of code - Emily (fixed issue where coming from wrong file)
os17.lung.raw <- tenXLoadQC(path10x = "/gpfs0/home2/gdrobertslab/lab/Counts/S0024xS0029/outs/filtered_feature_bc_matrix",
                            spec = "mixHuman")

os17.lung.raw <- subset(os17.lung.raw,
                        subset = nFeature_RNA > 1250 &
                          nCount_RNA < 60000 &
                          percent.mt < 25)

os17.lung.raw$src <- "Lung"
os17.lung.raw$cond <- "OS17"
```

### Process OS17 culture for Figure 1

```{r Fig1a1_process}
os17.cx <- NormalizeData(os17.cx.raw) %>%
  FindVariableFeatures(selection.method = "vst") %>%
  ScaleData() %>%
  RunPCA(pc.genes = os17.cx@var.genes,
         npcs = 20) %>%
  RunUMAP(reduction = "pca",
          dims = 1:20) %>%
  FindNeighbors(reduction = "pca",
                dims = 1:20) %>%
  FindClusters(resolution = 0.3)

DimPlot(os17.cx,
        reduction = "umap",
        pt.size = 1,
        label = T) +
  coord_fixed() +
  scale_color_npg(alpha = 0.7)

# Regress out the effects of cell cycle on these tumor cells
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes
os17.cx <- CellCycleScoring(object = os17.cx,
                            s.features = s.genes,
                            g2m.features = g2m.genes,
                            set.ident = TRUE)
os17.cx <- ScaleData(object = os17.cx,
                     vars.to.regress = c("S.Score",
                                         "G2M.Score"),
                     features = rownames(x = os17.cx))

os17.cx <- RunPCA(os17.cx,
                  pc.genes = os17.cx@var.genes,
                  npcs = 20) %>%
  RunUMAP(reduction = "pca",
          dims = 1:20) %>%
  FindNeighbors(reduction = "pca",
                dims = 1:20) %>%
  FindClusters(resolution = 0.3)

# Find optimal clustering resolution
os17.cx.nC <- nRes(os17.cx, 
                   res = seq(from = 0.1,
                             to = 0.2,
                             by = 0.05))
plot <- pSil(os17.cx.nC,
             0.15)
plot

# With res = 0.15, cells in cluster 3 do not have any significantly different genes that are deferentially regulated
# Therefore, we decided to proceed with res = 0.1
os17.cx <- FindClusters(os17.cx,
                        resolution = 0.1)

save(os17.cx,
     file ="/gpfs0/home2/gdrobertslab/rsexf001/OSHetero2021/Data/os17.cx_CCR.RData")
```

### Add lineage tags to metadata for Figure 5

For lineage analysis in Figure 5.

```{r Fig5_tag, dependson='load_os17'}
# Plot histograms of lineage tracing tags in the Seurat objects
os17.cx.raw <- processLTBC(os17.cx.raw,
                      lt.loc = "/gpfs0/home/gdrobertslab/lab/Counts/S0016/outs/lt.fq",
                      cid.loc = "/gpfs0/home/gdrobertslab/lab/Counts/S0016/outs/cid.fq",
                      histogram = F,
                      relative = T)

os17.tib.raw <- processLTBC(os17.tib.raw,
                       lt.loc = "/gpfs0/home/gdrobertslab/lab/Counts/S0018xS0028/outs/lt.fq",
                       cid.loc = "/gpfs0/home/gdrobertslab/lab/Counts/S0018xS0028/outs/cid.fq",
                       histogram = F,
                       relative = T)

os17.lung.raw <- processLTBC(os17.lung.raw,
                        lt.loc = "/gpfs0/home2/gdrobertslab/lab/Counts/S0024xS0029/outs/lt.fq",
                        cid.loc = "/gpfs0/home2/gdrobertslab/lab/Counts/S0024xS0029/outs/cid.fq",
                        histogram = F,
                        relative = T)
```

### Create OS17 object for lineage analysis in Figure 5

```{r Fig5_process, dependson='Fig5_tag'}
# Subset all to 2800 cells in each condition
os17.cx.raw <- subset(os17.cx.raw,
                 cells = sample(Cells(os17.cx.raw),
                                2800))
os17.tib.raw <- subset(os17.tib.raw,
                  cells = sample(Cells(os17.tib.raw),
                                 2800))
os17.lung.raw <- subset(os17.lung.raw,
                   cells = sample(Cells(os17.lung.raw),
                                  2800))

# Merge into a single Seurat object
os17 <- merge(os17.cx.raw,
              y = c(os17.tib.raw,
                    os17.lung.raw),
              add.cell.ids = c("Culture",
                               "Tibia",
                               "Lung"),
              project = "LineageTracing")

# Process and cluster
os17 <- NormalizeData(os17) %>%
  FindVariableFeatures(selection.method = "vst") %>%
  ScaleData() %>%
  RunPCA(pc.genes = os.17@var.genes, npcs = 20) %>%
  RunUMAP(reduction = "pca", dims = 1:20) %>%
  FindNeighbors(reduction = "pca", dims = 1:20) %>%
  FindClusters(resolution = 0.3)

# CCR Attempt to regress out the effects of cell cycle on these tumor cells
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes

os17 <- CellCycleScoring(object = os17,
                         s.features = s.genes,
                         g2m.features = g2m.genes,
                         set.ident = TRUE)

os17 <- ScaleData(object = os17,
                  vars.to.regress = c("S.Score",
                                      "G2M.Score"),
                  features = rownames(x = os17))

os17 <- RunPCA(os17, pc.genes = os.17@var.genes, npcs = 20) %>%
  RunUMAP(reduction = "pca", dims = 1:20) %>%
  FindNeighbors(reduction = "pca", dims = 1:20) %>%
  FindClusters(resolution = 0.3)

DimPlot(os17,
        reduction = "umap",
        pt.size = 1,
        label = T) +
  coord_fixed() +
  ggtitle("OS17 Clusters") +
  scale_color_npg(alpha = 0.7)

# Plot the data
set.seed(100)
cell.ids <- sample(colnames(os17))

DimPlot(os17,
        reduction = "umap",
        group.by = "src",
        pt.size = 1,
        label = F,
        order = cell.ids) +
  coord_fixed() +
  ggtitle("OS17 by Source") +
  scale_color_npg()

################## Find the optimum resolution for clustering
os17.nC <- nRes(os17,
                res = seq(from = 0.2,
                          to = 0.3,
                          by = 0.01))

save(os17,
     file = "/home/gdrobertslab/rsexf001/OSHetero2021/Data/os17.RData")
```

## Load t143b

```{r load_t143b}
# t143b
## Culture
t143b.cx.raw <- tenXLoadQC(path10x = "/gpfs0/home/gdrobertslab/lab/Counts/S0017/outs/filtered_feature_bc_matrix/",
                           spec = "mixHuman")

t143b.cx.raw <- subset(t143b.cx.raw,
                       subset = nFeature_RNA > 4000 &
                         nCount_RNA < 74000 &
                         percent.mt < 17)

VlnPlot(t143b.cx.raw,
        features = c("nFeature_RNA",
                     "nCount_RNA",
                     "percent.mt"),
        ncol = 3)

t143b.cx.raw$src <- "Culture"
t143b.cx.raw$cond <- "t143b"

## Tibia
t143b.tib.raw <- tenXLoadQC(path10x = "/gpfs0/home/gdrobertslab/lab/Counts/S0052/outs/filtered_feature_bc_matrix/",
                            spec = "mixHuman")

t143b.tib.raw <- subset(t143b.tib.raw,
                        subset = nFeature_RNA > 1000 &
                          nCount_RNA < 30000 &
                          percent.mt < 22)
VlnPlot(t143b.tib.raw,
        features = c("nFeature_RNA",
                     "nCount_RNA",
                     "percent.mt"),
        ncol = 3)

t143b.tib.raw$src <- "Tibia"
t143b.tib.raw$cond <- "t143b"

# Lung
t143b.lung.raw <- tenXLoadQC(path10x = "/gpfs0/home/gdrobertslab/lab/Counts/S0019-143B-lung/outs/filtered_feature_bc_matrix/",
                             spec = "mixHuman")

t143b.lung.raw <- subset(t143b.lung.raw,
                         subset = nFeature_RNA > 1000 &
                           nCount_RNA < 30000 &
                           percent.mt < 22)

VlnPlot(t143b.lung.raw,
        features = c("nFeature_RNA",
                     "nCount_RNA",
                     "percent.mt"),
        ncol = 3)

t143b.lung.raw$src <- "Lung"
t143b.lung.raw$cond <- "t143b"
```

## Load NCH-OS2

```{r load_nchos2}
# NCHOS2
## Flank
OS2.cx.raw <- tenXLoadQC(path10x = "/gpfs0/home/gdrobertslab/lab/Counts/S0076/outs/filtered_feature_bc_matrix/",
                         spec = "mixHuman")

OS2.cx.raw <- subset(OS2.cx.raw,
                     subset = nCount_RNA < 36000 &
                       percent.mt < 50)

VlnPlot(OS2.cx.raw,
        features = c("nFeature_RNA",
                     "nCount_RNA",
                     "percent.mt"),
        ncol = 3)

OS2.cx.raw$src <- "Flank"
OS2.cx.raw$cond <- "NCHOS2"

## Tibia
OS2.tib.raw<- tenXLoadQC(path10x = "/gpfs0/home/gdrobertslab/lab/Counts/S0042/outs/filtered_feature_bc_matrix/",
                         spec = "mixHuman")

OS2.tib.raw <- subset(OS2.tib.raw,
                      subset = nFeature_RNA > 2500 &
                        nCount_RNA < 40000 &
                        percent.mt < 20)

VlnPlot(OS2.tib.raw,
        features = c("nFeature_RNA",
                     "nCount_RNA",
                     "percent.mt"),
        ncol = 3)

OS2.tib.raw$src <- "Tibia"
OS2.tib.raw$cond <- "NCHOS2"

## Lung
OS2.lung.raw<- tenXLoadQC(path10x = "/gpfs0/home/gdrobertslab/lab/Counts/S0041/outs/filtered_feature_bc_matrix/",
                          spec = "mixHuman")
OS2.lung.raw <- subset(OS2.lung.raw,
                       subset = nFeature_RNA > 2500 &
                         nCount_RNA < 60000 &
                         percent.mt < 30)

VlnPlot(OS2.lung.raw,
        features = c("nFeature_RNA",
                     "nCount_RNA",
                     "percent.mt"),
        ncol = 3)

OS2.lung.raw$src <- "Lung"
OS2.lung.raw$cond <- "NCHOS2"
```

## Load NCH-OS7

```{r load_nchos7}
# NCHOS7
## Flank
OS7.cx.raw <- tenXLoadQC(path10x = "/gpfs0/home/gdrobertslab/lab/Counts/S0043/outs/filtered_feature_bc_matrix/",
                         spec = "mixHuman")

OS7.cx.raw <- subset(OS7.cx.raw,
                     subset = nCount_RNA < 50000 &
                       percent.mt < 25)

VlnPlot(OS7.cx.raw,
        features = c("nFeature_RNA",
                     "nCount_RNA",
                     "percent.mt"),
        ncol = 3)

OS7.cx.raw$src <- "Flank"
OS7.cx.raw$cond <- "NCHOS7"

## Tibia
OS7.tib.raw <- tenXLoadQC(path10x = "/gpfs0/home/gdrobertslab/lab/Counts/S0034/outs/filtered_feature_bc_matrix/",
                          spec = "mixHuman")
OS7.tib.raw <- subset(OS7.tib.raw,
                      subset = nFeature_RNA > 2000 &
                        nCount_RNA < 35000 &
                        percent.mt < 14)

VlnPlot(OS7.tib.raw,
        features = c("nFeature_RNA",
                     "nCount_RNA",
                     "percent.mt"),
        ncol = 3)

OS7.tib.raw$src <- "Tibia"
OS7.tib.raw$cond <- "NCHOS7"

## Lung
OS7.lung.raw <- tenXLoadQC(path10x = "/gpfs0/home/gdrobertslab/lab/Counts/S0055/outs/filtered_feature_bc_matrix/",
                           spec = "mixHuman")
OS7.lung.raw <- subset(OS7.lung.raw,
                       subset = nCount_RNA < 42000 &
                         percent.mt < 20)

VlnPlot(OS7.lung.raw,
        features = c("nFeature_RNA",
                     "nCount_RNA",
                     "percent.mt"),
        ncol = 3)

OS7.lung.raw$src <- "Lung"
OS7.lung.raw$cond <- "NCHOS7"
```

### Process NCHS-OS7 flank for Figure 1

```{r Fig1a3_process}
OS7.cx <- NormalizeData(OS7.cx.raw) %>%
  FindVariableFeatures(selection.method = "vst") %>%
  ScaleData() %>%
  RunPCA(pc.genes = OS7.cx@var.genes,
         npcs = 20) %>%
  RunUMAP(reduction = "pca",
          dims = 1:20) %>%
  FindNeighbors(reduction = "pca",
                dims = 1:20) %>%
  FindClusters(resolution = 0.3)

DimPlot(OS7.cx,
        reduction = "umap",
        pt.size = 1,
        label = T) +
  coord_fixed() +
  scale_color_npg(alpha = 0.7)

# Attempt to regress out the effects of cell cycle on these tumor cells
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes
OS7.cx <- CellCycleScoring(object = OS7.cx,
                               s.features = s.genes,
                               g2m.features = g2m.genes,
                               set.ident = TRUE)
OS7.cx <- ScaleData(object = OS7.cx,
                        vars.to.regress = c("S.Score",
                                            "G2M.Score"),
                        features = rownames(x = OS7.cx))

OS7.cx <- RunPCA(OS7.cx,
                     pc.genes = OS7.cx@var.genes,
                     npcs = 20) %>%
  RunUMAP(reduction = "pca",
          dims = 1:20) %>%
  FindNeighbors(reduction = "pca",
                dims = 1:20) %>%
  FindClusters(resolution = 0.3)

#Find optimal clustering resolution
OS7.cx.nC <- nRes(OS7.cx, 
                     res = seq(from = 0.1,
                               to = 0.15,
                               by = 0.01))
plot <- pSil(OS7.cx.nC,
             0.15)
plot

OS7.cx <- FindClusters(OS7.cx,
                          resolution = 0.15)

save(OS7.cx,
     file ="/gpfs0/home2/gdrobertslab/rsexf001/OSHetero2021/Data/OS7.cx_CCR.RData")
```


## Save Raw Objects

```{r save_raw}
# Save raw objects

# Make lists for easy loop saving
sample_list <- c(OB,
                 os17.cx.raw,
                 os17.tib.raw,
                 os17.lung.raw,
                 t143b.cx.raw,
                 t143b.tib.raw,
                 t143b.lung.raw,
                 OS2.cx.raw,
                 OS2.tib.raw,
                 OS2.lung.raw,
                 OS7.cx.raw,
                 OS7.tib.raw,
                 OS7.lung.raw)

sample_names <- c("OB",
                  "os17.cx.raw",
                  "os17.tib.raw",
                  "os17.lung.raw",
                  "t143b.cx.raw",
                  "t143b.tib.raw",
                  "t143b.lung.raw",
                  "OS2.cx.raw",
                  "OS2.tib.raw",
                  "OS2.lung.raw",
                  "OS7.cx.raw",
                  "OS7.tib.raw",
                  "OS7.lung.raw")

# Correlate sample names and sample labels
names(sample_list) <- sample_names

# Save each raw object
for (item in sample_names) {
  # Save each sample as an individual Seurat object with proper name
  assign(item, sample_list[[item]])
  save(list = item, file = paste("/gpfs0/home2/gdrobertslab/rsexf001/OSHetero2021/Data/Raw/",
                    item,
                    ".RData",
                    sep = ""))
}
```

# Merge Objects

Merge objects, process, and cell cycle regress. 

```{r mergeallobj}
# Merge into a single Seurat object
OS <- merge(OB, y = c(os17.cx.raw,
                      os17.tib.raw,
                      os17.lung.raw,
                      t143b.cx.raw,
                      t143b.tib.raw,
                      t143b.lung.raw,
                      OS2.cx.raw,
                      OS2.tib.raw,
                      OS2.lung.raw,
                      OS7.cx.raw,
                      OS7.tib.raw,
                      OS7.lung.raw),
            add.cell.ids = c("OB",
                             "OS17_Culture",
                             "OS17_Tibia",
                             "OS17_Lung",
                             "t143b_Culture",
                             "t143b_Tibia",
                             "t143b_Lung",
                             "NCHOS2_Flank",
                             "NCHOS2_Tibia",
                             "NCHOS2_Lung",
                             "NCHOS7_Flank",
                             "NCHOS7_Tibia",
                             "NCHOS7_Lung"),
            project = "Heterogeneity")

# Process and cluster
OS <- NormalizeData(OS) %>%
  FindVariableFeatures(selection.method = "vst") %>%
  ScaleData() %>%
  RunPCA(pc.genes = os.17@var.genes, npcs = 20) %>%
  RunUMAP(reduction = "pca", dims = 1:20) %>%
  FindNeighbors(reduction = "pca", dims = 1:20) %>%
  FindClusters(resolution = 0.3)

# rm(OB, os17.cx.raw, os17.tib.raw, os17.lung.raw, t143b.cx.raw, t143b.tib.raw,
# t143b.lung.raw, OS2.cx.raw, OS2.tib.raw, OS2.lung.raw, OS7.cx.raw, OS7.tib.raw,
# OS7.lung.raw)

# CCR Attempt to regress out the effects of cell cycle on these tumor cells
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes

OS <- CellCycleScoring(object = OS,
                       s.features = s.genes,
                       g2m.features = g2m.genes,
                       set.ident = TRUE)

OS <- ScaleData(object = OS,
                vars.to.regress = c("S.Score",
                                    "G2M.Score"),
                features = rownames(x = OS))

OS <- RunPCA(OS, pc.genes = OS@var.genes, npcs = 20) %>%
             RunUMAP(reduction = "pca", dims = 1:20) %>%
             FindNeighbors(reduction = "pca", dims = 1:20) %>%
             FindClusters(resolution = 0.3)

save(OS,
     file = "/gpfs0/home2/gdrobertslab/rsexf001/OSHetero2021/Data/OS_merged_postCCR.RData")

OS$cond <- as.factor(OS$cond)
OS$cond <- factor(as.factor(OS$cond),
                  levels = c("OS17",
                             "t143b",
                             "NCHOS2",
                             "NCHOS7",
                             "OB"))

# Plot the data
DimPlot(OS,
        reduction = "umap",
        group.by = "cond",
        pt.size = 1,
        label = T) +
  coord_fixed() +
  ggtitle("OS by Condition") +
    scale_color_npg(alpha = 1)

DimPlot(OS,
        reduction = "umap",
        group.by = "src",
        pt.size = 1,
        label = T) +
  coord_fixed() +
  ggtitle("OS by Source")

DimPlot(OS,
        reduction = "umap",
        pt.size = 1,
        label = T) +
  coord_fixed() +
  ggtitle("OS by Clusters") +
    scale_color_npg(alpha = 0.7)
```

# List Objects and Process

List objects, process, and cell cycle regress. Used in Figure 3 code.

```{r listallobj}
#Create List to compute ITH scores
OS.list <- list(OB = OB,
                OS17.cx = os17.cx.raw,
                OS17.Tibia = os17.tib.raw, 
                OS17.Lung = os17.lung.raw, 
                t143B.cx = t143b.cx.raw,
                t143B.Tibia = t143b.tib.raw, 
                t143B.Lung = t143b.lung.raw,
                OS2.Flank = OS2.cx.raw,
                OS2.Tibia = OS2.tib.raw, 
                OS2.Lung = OS2.lung.raw, 
                OS7.Flank= OS7.cx.raw,
                OS7.Tibia = OS7.tib.raw, 
                OS7.Lung = OS7.lung.raw) 

for (i in 1: length(OS.list)) {
  OS.list[[i]] <- NormalizeData(OS.list[[i]]) %>%
    FindVariableFeatures(selection.method = "vst") %>%
    ScaleData() %>%
    RunPCA(pc.genes = OS.list[[i]]@var.genes, npcs = 20) %>%
    RunUMAP(reduction = "pca", dims = 1:20) %>%
    FindNeighbors(reduction = "pca", dims = 1:20) %>%
    FindClusters(resolution = 0.3) %>%
    subset(cells = sample(Cells(OS.list[[i]]), 1500))
}

# CCR
# Attempt to regress out the effects of cell cycle on these tumor cells
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes

for (i in 1: length(OS.list)) {
  OS.list[[i]] <- CellCycleScoring(object = OS.list[[i]],
                                   s.features = s.genes,
                                   g2m.features = g2m.genes,
                                   set.ident = TRUE)
  
  OS.list[[i]] <- ScaleData(object = OS.list[[i]],
                            vars.to.regress = c("S.Score",
                                                "G2M.Score"),
                            features = rownames(x = OS.list[[i]]))
  
  OS.list[[i]] <- RunPCA(OS.list[[i]],
                         pc.genes = OS.list[[i]]@var.genes, npcs = 20) %>%
    RunUMAP(reduction = "pca", dims = 1:20) %>%
    FindNeighbors(reduction = "pca", dims = 1:20) %>%
    FindClusters(resolution = 0.3)
}

save(OS.list,
     file = "Data/OS.listv3.CCR.RData")
```

```{r versions}
sessionInfo()
```